---
title: "Metabolites_Paper_tab_Fig_19_Nov_24_final"
author: "Triet Thai"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE)

library(dplyr)
library(MASS)
library(parallel)
library(dplyr)
library(readxl)
library(writexl)
library(survival)
library(glmnet)
library(stabs)
library(tidyverse)
library(magrittr)
library(Hmisc)
library(gtsummary)
library(WGCNA)
library(ggcorrplot)
library(scales)
library(gridExtra)
library(ggcorrplot)
library(broom)
library(ggpubr)

source("Rfunctions/glmnet.R")
source("Rfunctions/GRaFo.R")
source("Rfunctions/cforest.R")
source("Rfunctions/glm.R")
source("Rfunctions/additional_functions.R")

path_input  <- ""
path_output <- ""
path_code   <- ""

mcore<-1
```

## Introduction

This markdown file contains the code to generate figures in Metabolites Paper.

Main figures include:

-   **Figure 1A: Relationship between metabolite abundance (log2 transformed) and estimated mortality (logisticcurve)**
-   **Figure 1B: Correlation plot for the 10 identified metabolites with Spearman's correlations using hierarchical clustering**
-   **Figure 1C: Violin plots show the distribution for the identified metabolites per patient group: TBM, CM, BM, and NIC.**
-   **Figure 2A: Stability variable selection through multivariable analysis**
-   **Figure 2B: The scatter plot visualized TBM patients following their log2-transformed CSF tryptophan and FA 8:0; 3OH concentrations, colored according to their day 60 outcome**
-   Figure 2C: Causal diagram showing the hypothesized relationship between CSF metabolites, TBM severity and mortality
-   **Figure 2D: Forest plots visualizing the estimates of direct and indirect impact of all top-hit metabolites on mortality**
-   Figure 3: Log2 fold changes for TBM patients vs. and non-infectious controls (NIC) in CSF and serum for top-hits following a previous cohort (van Laarhoven et al., 2018).
-   **Figure 4: Metabolite clustering and correlation with clinical characteristics.**
-   Figure 5: Diagram on dysregulated neuronal beta-oxidation as a source of the hydroxylated fatty acids that were associated to increased TBM mortality.

Supplementary figures include:

-   **Figure S1A: Spearman correlations and clustering of pre-treatment CSF metabolites**
-   **Figure S1B: Metabolite abundance ratio of TBM vs. non-infectious control**
-   **Figure S1C: Principal component projection of metabolite profiles in patients with TBM, bacterial meningitis, cryptococcal meningitis, and non-infectious controls**
-   **Figure S1D: Bi-directional ranking metabolites with regard to the association with two month mortality**
-   **Figure S1E: Ranking and forest plot of top-ranking metabolites.**
-   Figure S2.1: Abundances of available fatty acids in CSF and serum of TBM patients and non-infectious controls (NIC) and Spearman correlations with clinical parameters
-   Figure S2.2: Abundances of available carnitines in CSF and serum of TBM patients and non-infectious controls (NIC) and Spearman correlations with clinical parameters
-   Figure S3: Analysis workflow diagram
-   Figure S4: LC-MS-based analytical platform

## Main figures

### Figure 1: Pre-treatment metabolites and TBM two-month mortality and comparison with other patient groups

#### Figure 1A:

```{r}
library(cowplot)
library(dplyr)

Metabolites <- Metabol_tophit_names_combined
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="Unknown_of_interest","Unknown compound",names_ID$Metabolite)
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="2-hydroxy-3-methylpentanoate/hydroxyisocaproate","hydroxyisocaproate",names_ID$Metabolite)
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="alpha-hydroxybutyrate/beta-hydroxybutyrate/hydroxyisobutyrate","hydroxyisobutyrate",names_ID$Metabolite)

dat_top_hit_validate_VN <- subset(TBM_dat_VN,select=c(Metabolites,"age","HIV","ttdeath_2m","evdeath_2m","cohort"))
dat_top_hit_validate_IN <- subset(TBM_dat_IN,select=c(Metabolites,"age","HIV","ttdeath_2m","evdeath_2m","cohort"))

names(dat_top_hit_validate_VN) <- plyr::mapvalues(names(dat_top_hit_validate_VN), from = names_ID$Metabol_names, to = names_ID$Metabolite)
names(dat_top_hit_validate_IN) <- plyr::mapvalues(names(dat_top_hit_validate_IN), from = names_ID$Metabol_names, to = names_ID$Metabolite)
Metabolites <- plyr::mapvalues(Metabolites, from = names_ID$Metabol_names, to = names_ID$Metabolite)

dat_top_hit_validate_VN_2 <- subset(dat_top_hit_validate_VN,select=c(Metabolites,"age","HIV","ttdeath_2m","evdeath_2m"))
dat_top_hit_validate_IN_2 <- subset(dat_top_hit_validate_IN,select=c(Metabolites,"age","HIV","ttdeath_2m","evdeath_2m"))

d2<-rbind(dat_top_hit_validate_VN,dat_top_hit_validate_IN)
d2 %<>% filter(!(evdeath_2m==0&ttdeath_2m<60)) %>% dplyr::select(c(Metabolites,"evdeath_2m","cohort"))

d2_long <- gather(d2, metabolites, concentration, `N-carbamoyl-beta-alanine`:sebacate, factor_key=TRUE)
Get.breaks <- function(f, cuts, digits = 2)
{
  x <- round(quantile(f, probs = seq(1/cuts, 1 - 1/cuts, 1/cuts), names = F,na.rm=T), digits)
  x <- sort(unique(c(0, x, Inf)))
  rm(f, cuts, digits)
  return(x)
}

d2_long %<>% mutate(
  metabolites = dplyr::case_when(metabolites == "Unknown compound"~"unidentified compound",
                          T~metabolites)
)

```

```{r}
# GLOBAL PARAMs
Metabolites <- c(
  "hydroxyisocaproate",
  "3-hydroxyoctanoate",
  "hydroxyisobutyrate",
  "C4-OH carnitine",
  "tryptophan",
  "p-hydroxyphenylacetate",
  "phenyllactate",
  "N-carbamoyl-beta-alanine",
  "N6,N6,N6-trimethyllysine",
  "sebacate")

#original
rename_metabolites <- list("hydroxyisocaproate" = "FA 6:0;OH",
                           "3-hydroxyoctanoate"= "FA 8:0;3OH",
                           "hydroxyisobutyrate" = "FA 4:0;OH",
                           "C4-OH carnitine" = "CAR 4:0;OH",
                           "tryptophan" = "Tryptophan",
                           "p-hydroxyphenylacetate"="p-Hydroxyphenylacetate",
                           "phenyllactate" = "Phenyllactate",
                           "N-carbamoyl-beta-alanine" = "N-carbamoyl-beta-alanine",
                           "N6,N6,N6-trimethyllysine" = "N6,N6,N6-Trimethyllysine",
                           "sebacate" = "Sebacate")

dim(d2_long)#11594     4
cut_all = NULL
p<-list()
for(i in 1:length(Metabolites)) {
  df <-subset(d2_long,metabolites==Metabolites[i])
  breaks<-Get.breaks(df$concentration, cuts = 15, digits = 4)
  labels<-c(head(tail(breaks,-1),-1),mean(c(max(df$concentration,na.rm=T),head(tail(breaks,2),1))))
  concentration_max=max(df$concentration,na.rm=T)
  concentration_min=min(df$concentration,na.rm=T)
  cut <- df %>% mutate(metabolites_cut = cut(concentration, right = F,
                                             breaks = breaks,
                                             labels = labels),
                       concentration = as.numeric(as.character(metabolites_cut))) %>% 
        group_by(concentration,cohort) %>% 
        dplyr::summarise(.groups = "drop",
                         p_metabolites_cut = mean(evdeath_2m),
                         n_metabolites_cut= dplyr::n()) %>%
        mutate(se=sqrt(p_metabolites_cut*(1-p_metabolites_cut)/n_metabolites_cut)) %>%
        mutate(plo=p_metabolites_cut-qnorm(.975,0,1)*se, 
               phi=p_metabolites_cut+qnorm(.975,0,1)*se,
               concentration2=ifelse(cohort=="Vietnam",concentration+0.05,concentration)) %>% as.data.frame()

  cut$metabolites <- Metabolites[[i]]
  if(is.null(cut_all)){
    cut_all = cut
  }
  else {
    cut_all = rbind(cut_all,cut)
  }
  
  p_tmp<-ggplot(data=df, aes(x=concentration,y = evdeath_2m, color=cohort))+
          geom_smooth(method = glm,
                      aes(fill = cohort,group=cohort),
                      method.args = list(family = "binomial"),
                      se = TRUE,
                      alpha = 0.1,
                      size=2,fullrange=TRUE)+
          geom_point(data=cut,aes(x=concentration ,y=p_metabolites_cut,colour=cohort ),size=2.5,alpha=0.5)+
          geom_errorbar(data=cut,aes(x=concentration,
                                     y=p_metabolites_cut,
                                     ymin=plo,ymax=phi,
                                     col=cohort),
                        size=1,alpha=0.5,
                        )+
          coord_cartesian(xlim=c(concentration_min, concentration_max),ylim=c(0,1))+
          scale_x_continuous(breaks=breaks_pretty(n=4))+
          scale_color_manual(values=c("#313695", "#A50026"))+
          scale_fill_manual(values = c("#313695", "#A50026"))+
          labs(x = "",
               y = "",
               title = rename_metabolites[[Metabolites[i]]])+
          guides(fill=guide_legend(title="Cohort"),
                 color=guide_legend(title="Cohort", override.aes = list(size=14)))+
          theme_bw()+
          theme(plot.title = element_text(size = 25,hjust=0.5,face="bold"),
                axis.title.x = element_blank(),
                axis.title.y = element_blank(),
                axis.text.y = if(mod(i,5)==1) element_text( colour="black", size=25) else element_blank(),
                axis.text.x = element_text(colour="black", size=25),
                axis.ticks.y = if(mod(i,5)==1) element_line() else element_blank(),
                legend.title = element_text( colour="black", size=35,face="bold"),
                legend.text = element_text(size = 25),
                legend.key = element_rect(colour = NA),
                strip.text=element_text(face = "bold",size=15),
                plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
                panel.grid.minor = element_blank())
  
  leg <- ggpubr::get_legend(p_tmp)
  p_leg <- ggpubr::as_ggplot(leg)
  p_tmp <- p_tmp + theme(legend.position="none")
  
  xdens <- axis_canvas(p_tmp, axis = "x")+geom_density(data=df,aes(x=concentration,fill=cohort),alpha=0.25,size=0.5)+
           scale_fill_manual( values = c("#313695", "#A50026"))
  
  p1 <- insert_xaxis_grob(p_tmp, xdens, grid::unit(.2, "null"), position = "top")
  p[[i]]<-ggdraw(p1)
}

length(p)## 10

figure <- ggpubr::ggarrange(p[[1]],p[[2]],
                            p[[3]],p[[4]],
                            p[[5]],p[[6]],
                            p[[7]],p[[8]],
                            p[[9]],p[[10]],
                            widths = c(1.17,1,1,1,1),ncol =5,nrow=2) + theme(plot.margin = margin(1,1,1,1, "cm")) 
figure <- annotate_figure(figure,
                          left = text_grob("Sixty-day mortality", 
                                           color = "black", rot = 90, size = 30,face="bold",vjust=0.5),
                          bottom = text_grob("Intensity of ion peak", 
                                             color = "black", rot = 0, size = 30,face="bold",vjust = -0.2))
figure <- ggarrange(figure, p_leg,ncol=2,nrow=1,widths = c(9, 1))

# Export
tiff(file=paste0(path_output,"/Figure_1A.tiff"),res=1000, units = 'in', width=30, height =13,compression = 'lzw')
figure
dev.off()

```

#### Figure 1B:

```{r}

library(grid)
library(dplyr)
library(ggdendro)
library(stats)
library(ComplexHeatmap)
library(circlize)
library(dendextend)

cor.pval.test <- function(x1, x2, ...) {
  r <- ncol(x1)
  n <- ncol(x2)
  p.mat<- matrix(NA, r, n)
  for (i in 1:r) {
    for (j in 1:n) {
      tmp <- cor.test(x1[, i], x2[, j], ...)
      p.mat[i, j] <- tmp$p.value
    }
  }
  rownames(p.mat) <-  colnames(x1)
  colnames(p.mat) <- colnames(x2)
  return(p.mat)
}

data_pooled <- rbind(TBM_dat_IN,TBM_dat_VN) %>% mutate(TBM_grade_ord=as.numeric(factor(TBM_grade)),
                                                       TBM_grade_ord=ordered(TBM_grade_ord,levels=1:3))

names(data_pooled) <- plyr::mapvalues(names(data_pooled),
                                      from = names_ID$Metabol_names,
                                      to = names_ID$Metabolite)

df<-data_pooled %>% plotly::rename("CSF/blood glucose ratio"="CSF_blood_glucose_ratio","CSF lactate"="lactate",
                                   "CSF leukocytes counts" ="CSF_leukocytes",
                                   "CSF mononuclear cell counts" ="CSF_mononuclear_cells",
                                   "CSF polymorphonuclear cell counts" ="CSF_polymorphonuclear_cells",
                                   "CSF protein"="CSF_protein")

trp_mets <- c("sebacate",
              "N6,N6,N6-trimethyllysine",
              "N-carbamoyl-beta-alanine",
              "phenyllactate",
              "p-hydroxyphenylacetate",
              "tryptophan",
              "C4-OH carnitine",
              "hydroxyisobutyrate",
              "3-hydroxyoctanoate",
              "hydroxyisocaproate")

cor_method="spearman"
cor_mat <- cor(df[, trp_mets],use='pairwise.complete.obs', method = cor_method)
p.cor_mat <- cor.pval.test(df[, trp_mets], df[, trp_mets],method=cor_method, use='pairwise.complete.obs')

dist_matrix <- dist(abs(cor_mat), method="euclidean")
hc <- hclust(dist_matrix, method="ward.D")
hcd <- as.dendrogram(hc)

mat_tmp <- cor_mat
p.mat_tmp <- p.cor_mat
labels1 <-round(mat_tmp,2)
vals1 <- labels1
vals1[p.mat_tmp > 0.05] <- 0
labels1[p.mat_tmp > 0.05] <- ''


col=c("#001164", 'white', "#FF681E")
col_fun = colorRamp2(c(-1,0,1), col)

tmp0 <- rev(hcd)
tmp <- rev(hcd)
order.dendrogram(tmp) <- c( 10,9,8,7,6,5,4,3,2,1)
tmp0 <- match_order_by_labels(tmp0, tmp)

colnames(vals1) <- rev(unname(unlist((rename_metabolites))))
rownames(vals1) <- rev(unname(unlist((rename_metabolites))))

#Heatmap
row_ha = rowAnnotation(month = anno_text(colnames(vals1), just = "center", gp = gpar(fontsize = 14,fontface='bold'),
        location = unit(0.5, "npc"), show_name = FALSE), annotation_name_rot = 0)

hm = Heatmap(vals1,col = col_fun,
        cell_fun = function(j, i, x, y, width, height, fill) {
        grid.text(labels1[i,j], x, y, gp = gpar(fontsize = 10))},
        rect_gp = gpar(col = "black", lwd = 0.5),
        row_order = colnames(vals1),
        cluster_rows = tmp0,
        cluster_columns = tmp0,
        row_names_side = "right",
        row_dend_side = "left",
        show_heatmap_legend = FALSE,
        name = "Corr.",
        show_row_names = TRUE,
        column_names_rot = 45,
        column_names_gp = gpar(fontsize = 14,fontface='bold'),
        row_names_gp = gpar(fontsize = 14,fontface='bold'),
        row_dend_width = unit(2, "cm"),
        column_dend_height = unit(2, "cm"),)

lgd = Legend(col_fun = col_fun, title = "Corr.", legend_height = unit(3, "in"), at = c(-1, -0.5,0, 0.5, 1),
             labels_gp = gpar(col = "black", fontsize = 14), title_gp = gpar(col = "black", fontsize = 14,fontface ='bold'))

p <- draw(hm, heatmap_legend_list = lgd, padding = unit(c(2, 2, 2, 2), "mm"))


#Export
tiff(file=paste0(path_output,"/Figure_1B.tiff"),res=480, units = 'in', width = 7.5, height =6.5,compression = 'lzw')
draw(hm, heatmap_legend_list = lgd, padding = unit(c(2, 2, 2, 2), "mm"))
dev.off()
```

#### Figure 1C:

```{r}
library(cowplot)
library(ggpubr)
library(rstatix)
library(tidyverse)
library(cowplot)
library(DescTools)
library(ggsci)


Metabolites <- Metabol_tophit_names_combined
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="Unknown_of_interest","Unknown compound",names_ID$Metabolite)
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="2-hydroxy-3-methylpentanoate/hydroxyisocaproate","hydroxyisocaproate",names_ID$Metabolite)
names_ID$Metabolite<-ifelse(names_ID$Metabolite=="alpha-hydroxybutyrate/beta-hydroxybutyrate/hydroxyisobutyrate","hydroxyisobutyrate",names_ID$Metabolite)

dat_top_hit_all <- subset(clinmet_full,select=c(Metabolites,"age","HIV","ttdeath_2m","evdeath_2m","diagnosis_ultimate2"))

names(dat_top_hit_all) <- plyr::mapvalues(names(dat_top_hit_all), from = names_ID$Metabol_names, to = names_ID$Metabolite)
Metabolites <- plyr::mapvalues(Metabolites, from = names_ID$Metabol_names, to = names_ID$Metabolite)

dat_top_hit_all_2 <- subset(dat_top_hit_all,select=c(Metabolites,"ttdeath_2m","evdeath_2m","diagnosis_ultimate2"))

# d2<-dat_top_hit_all_2 %>% mutate(diagnosis_ultimate3=ifelse(diagnosis_ultimate2%in%c("Bacterial meningitis", "Cryptococcal meningitis", ),"Other CNS infection",diagnosis_ultimate2))
d2<-dat_top_hit_all_2 %>% mutate(diagnosis_ultimate3=diagnosis_ultimate2)# all

d2 %<>% filter(!(evdeath_2m==0&ttdeath_2m<61)) %>% dplyr::select(c(Metabolites,"evdeath_2m","diagnosis_ultimate3"))

d2_long <- gather(d2, metabolites, concentration, `N-carbamoyl-beta-alanine`:sebacate, factor_key=TRUE)

dim(d2_long) #12320     4

d2_long %<>% mutate(etiology=ifelse(diagnosis_ultimate3=="TBM" & evdeath_2m==1,'TBM non survivors',
                                    ifelse(diagnosis_ultimate3=="TBM" & evdeath_2m==0,
                                           'TBM survivors', diagnosis_ultimate3)))

my_comparisons <- list( c("Bacterial meningitis","Non-infectious control"),
                        c("Cryptococcal meningitis","Non-infectious control"),
                        c("TBM survivors","Non-infectious control"),
                        c("TBM non survivors","Non-infectious control"))

Metabolites <- c(
  "hydroxyisocaproate",
  "3-hydroxyoctanoate",
  "hydroxyisobutyrate",
  "C4-OH carnitine",
  "tryptophan",
  "p-hydroxyphenylacetate",
  "phenyllactate",
  "N-carbamoyl-beta-alanine",
  "N6,N6,N6-trimethyllysine",
  "sebacate")

rename_metabolites <- list("hydroxyisocaproate" = "FA 6:0;OH",
                           "3-hydroxyoctanoate"= "FA 8:0;3OH",
                           "hydroxyisobutyrate" = "FA 4:0;OH",
                           "C4-OH carnitine" = "CAR 4:0;OH",
                           "tryptophan" = "Tryptophan",
                           "p-hydroxyphenylacetate"="p-Hydroxyphenylacetate",
                           "phenyllactate" = "Phenyllactate",
                           "N-carbamoyl-beta-alanine" = "N-carbamoyl-beta-alanine",
                           "N6,N6,N6-trimethyllysine" = "N6,N6,N6-Trimethyllysine",
                           "sebacate" = "Sebacate")

p<-list()
library(ggbeeswarm)
for(i in 1:length(Metabolites)){
  print(i)
  if(mod(i,5)==1){
    ticksy <- NULL
    texty <- element_text( colour="black", size=25)
    }
  else{
      ticksy <- element_blank()
      texty <- element_blank()
  }

  df <-subset(d2_long,metabolites==Metabolites[i])
  df$etiology <- factor(df$etiology,     # Reorder factor levels
                        c("Non-infectious control", "Bacterial meningitis", "Cryptococcal meningitis", "TBM survivors","TBM non survivors"))
  
  stat.test <- df %>%
  group_by('etiology') %>% 
  wilcox_test(concentration~etiology, p.adjust.method = 'none') %>%
  filter(group1=="Non-infectious control") %>%
  add_xy_position(x = 'etiology')
  
 p1<-ggplot(data=df,aes(y=concentration, x=etiology,fill=etiology,color=etiology)) +
   geom_quasirandom(alpha = 0.8,shape=21,size=4)+
   geom_boxplot(width=0.3, fill="white",col='black',alpha=0.65) +
   
   stat_pvalue_manual(inherit.aes = FALSE, stat.test, hide.ns = TRUE,step.increase=0.1,label.size=7)+
   scale_y_continuous(breaks=breaks_pretty(n=4))+
   
   scale_color_manual(values=c("#bebada", "#ebbb8b","#A3BE8C","#81A1C1","#926189")) + ##88C0D0
   scale_fill_manual(values=c("#bebada", "#ebbb8b","#A3BE8C","#81A1C1","#926189")) +
     
   labs(y = "",x = "",title = rename_metabolites[[Metabolites[i]]])+
   guides(fill = guide_legend(title="Patient category"),
          color = guide_legend(title="Patient category", override.aes = list(size=14)))+
   theme_bw()+
   theme(plot.title = element_text(size = 25,hjust=0.5,face="bold"),
         axis.title.x = element_text(face="bold", colour="black", size=25),
         axis.title.y = element_text(face="bold", colour="black", size=25),
         axis.text.y = element_text( colour="black", size=25),
         axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         legend.text = element_text(size = 25),
         legend.title = element_text( colour="black", size=35,face='bold'),
         legend.key = element_rect(colour = NA),
         legend.key.size = unit(1.5, 'cm'),
         panel.grid.minor = element_blank(),)
 leg <- ggpubr::get_legend(p1)
 p_leg <- ggpubr::as_ggplot(leg)
 p1 <- p1 + theme(legend.position="none")
 p[[i]]<-ggdraw(p1)
}

figure = ggpubr::ggarrange(p[[1]],p[[2]],
                           p[[3]],p[[4]],
                           p[[5]],p[[6]],
                           p[[7]],p[[8]],
                           p[[9]],p[[10]],
                           ncol = 5,nrow = 2)

figure <- annotate_figure(figure,left = text_grob("Metabolite concentration\n(log 2 of ion peak intensity)", color = "black", rot = 90, size = 35,face="bold",vjust=0.5)) #"Intensity of ion peak"


figure <- ggpubr::ggarrange(figure,NULL,p_leg,nrow=1,widths=c(5,0.05,1))

#export
tiff(file=paste0(path_output,"/Figure_1C.tiff"),res=480, units = 'in', width=32, height =10,compression = 'lzw')
figure
dev.off()

```

### Figure 2:

#### Figure 2A:

```{r}

#both In & VN
load(file="Derive_Results/variable_selection_in_both_IN_VN.Rdata")
dat_plot<-data.frame(variable=names(stab.boot$max),`Selected probability`=stab.boot$max)[-1,]
dat_plot$variable <- plyr::mapvalues(dat_plot$variable, from = c(names_ID$Metabol_names,"HIV_bi"), to = c(names_ID$Metabolite,"HIV"))

dat_plot <- dat_plot[!dat_plot$variable == "Unknown compound", ]

dat_plot$variable <- plyr::mapvalues(dat_plot$variable, from = c(names(rename_metabolites),'Ct','age'), to = c(unlist(unname(rename_metabolites)),'Bacterial load','Age'))

dat_plot %>%
  mutate(name = reorder(variable, Selected.probability)) %>%
  ggplot(aes(y=name,x=Selected.probability))+geom_point(col="#2c7fb8",size=4,shape=15)+
  geom_vline(xintercept=stab.boot$cutoff,col="red")+
  geom_line(linetype = "dashed")+
  xlab("Selected probability")+theme_bw()+theme(legend.position="bottom")+ theme(panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=20),
        axis.title.y = element_blank(),
        axis.text.y = element_text(face = "bold", colour="black", size=18),
        axis.text.x = element_text( colour="black", size=15),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=15),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=15),
        plot.title = element_text(size = 20, face = "bold",hjust = 0.5))->p1

tiff(file=paste0(path_output,"/Figure_2A.tiff"),res=480, units = 'in', width =9*1.1, height = 6*1.1,compression = 'lzw')
p1
dev.off()
#########################################################################################################################
# load(file="Derive_Results/variable_selection_in_VN.Rdata")
# dat_plot<-data.frame(variable=names(stab.boot$max),`Selected probability`=stab.boot$max)[-1,]
# dat_plot$variable <- plyr::mapvalues(dat_plot$variable, from = c(names_ID$Metabol_names,"HIV_bi"), to = c(names_ID$Metabolite,"HIV"))
# 
# dat_plot <- dat_plot[!dat_plot$variable == "Unknown compound", ]
# 
# dat_plot$variable <- plyr::mapvalues(dat_plot$variable, from = c(names(rename_metabolites),'Ct','age'), to = c(unlist(unname(rename_metabolites)),'Bacterial load','Age'))
# 
# dat_plot %>%
#   mutate(name = reorder(variable, Selected.probability)) %>%
#   ggplot(aes(y=name,x=Selected.probability))+geom_point(col="#2c7fb8",size=4,shape=15)+
#   geom_vline(xintercept=stab.boot$cutoff,col="red")+
#   geom_line(linetype = "dashed")+
#   xlab("Selected probability")+theme_bw()+theme(legend.position="bottom") + 
#   theme(panel.grid.minor = element_blank(),
#         axis.title.x = element_text(face="bold", colour="black",size=20),
#         axis.title.y = element_blank(),
#         axis.text.y = element_text(face = "bold", colour="black", size=18),
#         axis.text.x = element_text( colour="black", size=15),
#         axis.ticks.x=element_blank(),
#         legend.text = element_text(size = 15),
#         legend.title = element_text( colour="black", size=15),
#         legend.key = element_rect(colour = NA),
#         strip.text=element_text(face = "bold",size=15),
#         plot.title = element_text(size = 20, face = "bold",hjust = 0.5))->p2
```

#### Figure 2B:

```{r}

library(Matrix)
library(tidyverse)
library(magrittr)
library(dplyr)
library(patchwork)
library(ggplot2)
library(plyr)
library(lattice)
library(MASS)
library(viridis)
library(colorspace)
library(wesanderson)
library(ggsci)
library(scales)
library(NatParksPalettes)
library(MASS)
library(metR)
library(geomtextpath)
library(raster)
library(sp)
library(ggalt)
library(sf)
library(smoothr)
library(ggforce)
library(DescTools)
library(ggfx)
library(scattermore)
library(ggnewscale)

l <-load(paste0(path_code,"/clinment.Rdata"))
clinmet %<>% rename_all(~ gsub("_", "-", .))
data <- clinmet[names(clinmet)[1:470]] #metabolites
print(data[!complete.cases(data), ]) #5 rows missing value
missing_id <-data[!complete.cases(data), ]$PatientID
clinmet <- clinmet[!(clinmet$PatientID %in% missing_id), , drop = FALSE]

data <- clinmet[names(clinmet)[1:470]]
rownames(data) <- data$PatientID
data$PatientID <- NULL

feats <- c('QI04209_HILIC_pos', 'TF5_HILIC_neg', 'QI05742_C18_neg', 'QI06445_C18_neg', 'TF19_HILIC_pos', 'QI32_HILIC_neg', 
         'TF21_HILIC_pos', 'QI09136_HILIC_pos', 'QI05254_C18_neg')

feats <- lapply(X = feats, FUN = function(t) gsub(pattern = "_", replacement =  "-", x = t, fixed = TRUE))

clinmet <- clinmet %>% filter(!((clinmet['ttdeath-2m']<61) & (clinmet['evdeath-2m']==0)))

print("Data cleaning complete")
print(dim(data)) #1062  469

GCS_level <-cut(clinmet$GCS,
    breaks = c(4,11,15,16),
    include.lowest = TRUE,
    right = F)

clinmet$GCS_level <- mapvalues(GCS_level, from=c("[4,11)","[11,15)","[15,16]"), to=c("4-10","11-14","15"))

df <- clinmet[c("QI06445-C18-neg","TF19-HILIC-pos","evdeath-2m","GCS_level","GCS")] %>% 
      mutate(GCS = as.factor(GCS),`evdeath-2m` = as.factor(mapvalues(`evdeath-2m`, from=c("0","1"), to=c("alive","dead"))))
names(df) <- lapply(X = names(df), FUN = function(t) gsub(pattern = "-", replacement =  "_", x = t, fixed = TRUE))
```

```{r}
sf::sf_extSoftVersion()["GEOS"] #"3.11.2"
geos::geos_version() #> '3.11.1'

elp <- ggplot(data =df,aes(TF19_HILIC_pos,QI06445_C18_neg))+
              stat_density_2d(n=500,breaks=c(0.00043),geom='polygon')+
              xlim(c(2,15))+
              ylim(c(10,22))
elp <- ggplot_build(elp)$data[[1]]
hull <- elp %>% st_as_sf(coords=c("x", "y")) %>% st_union() %>% st_concave_hull(ratio = .01)
p <- as(hull, 'Spatial')
r <- rasterFromXYZ(x_tilde[,c('TF19_HILIC_pos','QI06445_C18_neg','mortality')])
rp <- raster::mask(r,p)
x_tilde2 <- as.data.frame(rp, xy = TRUE)

p <-ggplot() + 
    geom_raster(data = x_tilde2, mapping = aes(x=x,y = y, fill=mortality),interpolate = TRUE) +
    scale_fill_distiller(name = "Predicted mortality",palette = "RdYlBu",na.value="white") +
    geom_point(data=df%>%filter(evdeath_2m=='alive'), 
               aes(x=TF19_HILIC_pos,y= QI06445_C18_neg,color=factor(evdeath_2m)),
               size=5,alpha=0.4,show.legend = T) +
    geom_point(data=df%>%filter(evdeath_2m=='dead'), aes(x=TF19_HILIC_pos,
                                                         y= QI06445_C18_neg,
                                                         color=factor(evdeath_2m)),
               size=5, alpha=0.3, show.legend = T) +
    scale_color_manual(name = "Patient outcome",
                      values = c(alive = "#f79aaf", dead = "black"),
                      labels = c("alive", "dead"),
                      guide = guide_legend(override.aes = list(linetype = c( alive = "blank",dead = "blank"),
                                                               shape = c(alive = 16, dead = 16)))) +
    new_scale_color() + 
    geom_contour(data = x_tilde,aes(color='line1',x=TF19_HILIC_pos,y= QI06445_C18_neg,z=GCS),
                 bins=15,show.legend=T,breaks=c(11,12,13,14,15)) + 
    geom_text_contour(data= x_tilde,
                    aes(x=TF19_HILIC_pos,y= QI06445_C18_neg,z=GCS),
                    skip=0, breaks=c(11,12,13,14,15), stroke = 0.1, stroke.colour = "white",
                    label.placer = label_placer_fraction(frac = 0.5,rot_adjuster = isoband::angle_halfcircle_bottom()),
                    colour = "#006D5B",
                    fontface = "bold",
                    size=7,show.legend = F,) + 
    scale_colour_manual(name = "Severity",
                      values = c("line1" = "#006D5B"),
                      labels = c("GCS"),
                      guide = guide_legend(override.aes = list(linetype = c("line1" = "solid"),shape = c("line1" = NA)))) +
    labs(x="Tryptophan", y="FA 8:0;3OH", color='Severity',size="Predicted mortality", shape='Patient outcome') + 
    guides(color = guide_legend(order=1,override.aes = list(shape = NA,size=7)),
           size = guide_legend(order=2,override.aes = list(size=7)),
           shape = guide_legend(order=3,override.aes = list(size=7)))+
    theme_classic() +
    theme(panel.grid.minor = element_blank(),
          axis.title.x = element_text(face="bold", colour="black", size=25),
          axis.title.y = element_text(face="bold", colour="black", size=25),
          axis.text.y = element_text( colour="black", size=20),
          axis.text.x = element_text( colour="black", size=20),
          axis.ticks.x=element_blank(),
          legend.text = element_text(size = 20),
          legend.title = element_text( colour="black", size=25,face='bold'),
          legend.key = element_rect(colour = NA),
          strip.text=element_text(face = "bold",size=15))

p <- ggplot_build(p)
p <- ggplot_gtable(p)

tiff(file=paste0(paste0(path_output,"/Figure_2B.tiff")),res=480, units = 'in', width = 12, height = 8,compression = 'lzw')
grid::grid.draw(p)
dev.off()

```

#### Figure 2D:

```{r}
library(grid)

load(file="Derive_Results/mediation_metabolites.Rdata")
p_mediation<-data.table::rbindlist(p)

mediated_metabolites <- p_mediation %>% filter(Effect=="Rtnie") %>% dplyr::select(metabolite) %>% c()

order_vector <- rev(Metabolites)

p_mediated_long<-p_mediation %>% mutate(Effect=case_when(Effect=="Rtnde"~"Direct impact",
                                                         Effect=="Rtnie"~"Indirect impact",
                                                         Effect=="Rte"~"Total impact",
                                                         .default ="Proportion Mediated")) %>%
                                 mutate(Effect=factor(Effect,levels=c("Direct impact","Indirect impact",
                                                                      "Total impact","Proportion Mediated"))) 
  
dim(p_mediated_long)# 48  7
p_mediated_long_rm<-p_mediated_long %>% filter(metabolite != "Unknown coumpound") %>% 
                                        filter(metabolite != "3-hydroxyhexanoate")

p_mediated_long_rm %>% filter((Effect!="Proportion Mediated")&(metabolite%in%order_vector)) %>%
                       mutate(metabolite=factor(metabolite, levels = order_vector, ordered = TRUE)) %>% 
                       
  ggplot(aes(x=metabolite, y=Estimate,group=Effect)) +
  geom_point(color="#313695",size=2) + 
  facet_grid(~Effect,scales = "free_x")+
  coord_trans(y = "log2")+
  geom_errorbar(aes(ymin = `95% CIL`, ymax = `95% CIU`),width=.2,size=1,alpha=0.5,color="#313695")+
  coord_flip() +
  scale_x_discrete("",labels = rename_metabolites)+
  xlab("Metabolites") +
  ylab("Hazard ratio")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20),
    legend.position="bottom",
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(face="bold", colour="black", size=25),
        panel.spacing = unit(1, "lines"),
        axis.title.y = element_text(face="bold", colour="black", size=25),
        axis.text.y = element_text(face="bold", colour="black", size=18),
        axis.text.x = element_text(face="bold", colour="black", size=18),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=16),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=20)) -> p1

dim(p_mediated_long_rm)# 40  7
p_mediated_long_rm%>% filter((Effect=="Proportion Mediated")&(metabolite%in%mediated_metabolites$metabolite)) %>% 
  mutate(metabolite=factor(metabolite, levels = order_vector, ordered = TRUE)) %>% 
  ggplot(aes(x=metabolite, y=Estimate)) +
  geom_point(color="#A50026",size=2) +facet_grid(~Effect, scales="free_x")+
  geom_errorbar(aes(ymin = `95% CIL`, ymax = `95% CIU`),width=.2,size=1,alpha=0.5,color="#A50026")+
  coord_flip() +
  xlab("") +
  ylab("Proportion")+
  theme_bw()+
  ylim(c(-0.05,1))+
  theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20),
        panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=25),
        axis.title.y = element_text(face="bold",colour="black", size=25),
        axis.text.y = element_blank(),
        axis.text.x = element_text(face="bold",colour="black", size=18),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none",
        legend.text = element_text(size = 20),
        legend.title = element_text( colour="black", size=35),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=20))->p2

#Export
tiff(file=paste0(path_output,"/Figure_2D.tiff"),res=480, units = 'in', width = 18, height =6,compression = 'lzw')
grid.newpage()
grid.draw(cbind(ggplotGrob(p1),
                 ggplotGrob(p2),
                 size = "last"))
dev.off()
```

### Figure 4:

```{r}
library(grid)

data_pooled <- rbind(TBM_dat_IN,TBM_dat_VN) %>% mutate(TBM_grade_ord=as.numeric(factor(TBM_grade)),
                                                       TBM_grade_ord=ordered(TBM_grade_ord,levels=1:3))
names(data_pooled) <- plyr::mapvalues(names(data_pooled),
                                      from = names_ID$Metabol_names,
                                      to = names_ID$Metabolite)
dim(dat_IN) #388  28
dim(dat_VN) #679  63

characteristics_numeric1 <- c('age','GCS',
                             'CSF mononuclear cell counts',
                             'CSF polymorphonuclear cell counts',
                             'CSF protein',
                             "CSF/blood glucose ratio",'Ct')

characteristics_numeric2 <-  c("log2.TNFa","log2.IFNg","log2.IL2","log2.IL4","log2.IL5","log2.IL6","log2.IL10","log2.IL13")

trp_mets <- c("sebacate",
              "N6,N6,N6-trimethyllysine",
              "N-carbamoyl-beta-alanine",
              "phenyllactate",
              "p-hydroxyphenylacetate",
              "tryptophan",
              "C4-OH carnitine",
              "hydroxyisobutyrate",
              "3-hydroxyoctanoate",
              "hydroxyisocaproate")

col=c("#001164", 'white', "#FF681E")

# Pooled cohort
df1<-data_pooled %>% plotly::rename("CSF/blood glucose ratio"="CSF_blood_glucose_ratio","CSF lactate"="lactate",
                                   "CSF leukocytes counts" ="CSF_leukocytes",
                                   "CSF mononuclear cell counts" ="CSF_mononuclear_cells",
                                   "CSF polymorphonuclear cell counts" ="CSF_polymorphonuclear_cells",
                                   "CSF protein"="CSF_protein")

mat1 <- cor(df1[, trp_mets],df1[, characteristics_numeric1],
            use='pairwise.complete.obs', method = cor_method)

p.mat1 <- cor.pval.test(df1[, trp_mets],df1[, characteristics_numeric1],
                        method=cor_method, use='pairwise.complete.obs')

colnames(mat1) <- plyr::mapvalues(colnames(mat1),from = c('age','Ct'), to = c('Age','GeneXpert Ct-value'))
colnames(p.mat1) <- plyr::mapvalues(colnames(p.mat1),from = c('age', 'Ct'), to = c('Age', 'GeneXpert Ct-value'))

rownames(mat1) <- rev(unname(unlist((rename_metabolites))))[1:10]
rownames(p.mat1) <- rev(unname(unlist((rename_metabolites))))[1:10]

p1  <- ggcorrplot(t(mat1),
             p.mat = t(p.mat1), 
             colors = col,lab = T, lab_size = 4,insig="blank",
             outline.color = 'black')+ theme(legend.position="none")+
    labs(fill="Corr")+
    theme(axis.text.x = element_text(face="bold", colour="black", size=16),
          axis.text.y = element_text(face="bold", colour="black", size=16),
          legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=12)) + 
    theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20))+
  labs(title = element_text(hjust = 100,vjust=50, "Combined cohort",colour="black", size=12,face='bold'))

## Indonesia cohort
df2<-dat_IN
mat2 <- cor(df2[, characteristics_numeric2],df2[, trp_mets],
            use='pairwise.complete.obs', method = cor_method)

p.mat2 <- cor.pval.test(df2[, characteristics_numeric2],df2[, trp_mets ],
                          method=cor_method, use='pairwise.complete.obs')

renamed_chars <- gsub("log2.", "", characteristics_numeric2)

rownames(mat2) <- plyr::mapvalues(rownames(mat2),
                                  from = characteristics_numeric2, to = renamed_chars)
rownames(p.mat2) <- plyr::mapvalues(rownames(p.mat2),
                                    from = characteristics_numeric2, to = renamed_chars)

p2 <-ggcorrplot(mat2,p.mat = p.mat2, insig = 'blank',colors = col,
                outline.color = 'black',legend.title = "Corr.",
                lab = T, lab_size = 4)+ylab("")+xlab("")+
  theme(legend.position = 'none')+
  theme(panel.grid.minor = element_blank(),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(2, "cm"),
        axis.title.x = element_text(face="bold", colour="black", size=12),
        axis.title.y = element_text(face="bold", colour="black", size=12),
        axis.text.y = element_blank(),
        axis.text.x = element_text(face="bold", colour="black", size=16),
        axis.ticks.x= element_blank(),
        axis.ticks.y= element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=12),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=12),
        plot.margin=unit(c(-0.2,-0.1,-0.6,-0.1), "cm")) +
  theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20)) + 
  labs(title = element_text(hjust = 100,vjust=50, "Indonesia",colour="black", size=12,face='bold'))

#### Vietnam cohort
df3<-dat_VN
mat3 <- cor( df3[, characteristics_numeric2],df3[, trp_mets],
            use='pairwise.complete.obs', method = cor_method)

p.mat3 <- cor.pval.test(df3[, characteristics_numeric2],df3[, trp_mets],
                        use='pairwise.complete.obs',method=cor_method)

colnames(mat3) <- rev(unname(unlist((rename_metabolites))))
colnames(p.mat3) <- rev(unname(unlist((rename_metabolites))))

rownames(mat3) <- plyr::mapvalues(rownames(mat3),
                                  from = characteristics_numeric2, to = renamed_chars)
rownames(p.mat3) <- plyr::mapvalues(rownames(p.mat3),
                                    from = characteristics_numeric2, to = renamed_chars)

p3 <- ggcorrplot(mat3,
             p.mat = p.mat3, insig = 'blank',
             colors = col,
             outline.color = 'black',
             lab = T, lab_size = 4)+ylab("")+xlab("") +
  theme(legend.position = 'right')+
  theme(panel.grid.minor = element_blank(),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(1, "cm"),
        axis.title.x = element_text(face="bold", colour="black", size=12),
        axis.title.y = element_text(face="bold", colour="black", size=12),
        axis.text.y = element_blank(),
        axis.text.x = element_text(face="bold", colour="black", size=16),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=12,face='bold'),
        legend.key = element_rect(colour = NA))+
  theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20))+
  labs(title = element_text(hjust = 100,vjust=50, "Vietnam",colour="black", size=12,face='bold'))

# Merge and export
tiff(file=paste0(path_output,"/Figure_4.tiff"),res=900, units = 'in', width = 13*1.1, height =10*1.1,compression = 'lzw')
grid.newpage()
grid.draw(cbind(ggplotGrob(p1),
                ggplotGrob(p2),
                ggplotGrob(p3),
                size = "last"))
dev.off()

```

## Supplemetary Figures

### Figure S1A:

```{r}
library(pheatmap)
library(ggplotify)
library(grid)

clinmet <- clinmet_full

hmdb_annotation <- read.csv(paste0(path_input, '/hmdb_all_metabolites.csv'))
hmdb_annotation %>% 
  dplyr::select(accession, name, kegg, kingdom, direct_parent, super_class, class,
         sub_class) %>% 
  as.data.frame() -> hmdb_annotation

# List of duplicated metabolites (measured by more than one LC-MS platform)
dup_mets <- read_excel(paste0(path_input, '/0_Duplicate_metabolites_missing_count_selection_from_Julian_GC.xlsx'))
dup_mets_id <- dup_mets$row.names[dup_mets$selected == 0]

# List patient outliers
# outliers <- c('900526','900520')

#Remove duplicates and outliers
untargeted %>%
  filter(!(CompoundID %in% dup_mets_id)) %>%
  as.data.frame() -> untargeted_no_overlap

#Merge with annotation
untargeted_annotated <- merge(untargeted_no_overlap, hmdb_annotation, 
                              by.x = 'HMDB_ID', by.y = 'accession',
                              all.x = T)

met_annotation <- untargeted_annotated %>%
  dplyr::select(CompoundID, super_class) %>%
  tibble::column_to_rownames('CompoundID') %>%
  as.data.frame()
rownames(met_annotation)<-gsub("-", "_", rownames(met_annotation), fixed = TRUE)
met_annotation$super_class[is.na(met_annotation$super_class)] <- 'Others'

paletteLength <- 50
#myColor <- colorRampPalette(c("#313695", 'white', "#A50026"))(paletteLength)
myColor <- colorRampPalette(c("#001164", "white", "#FF681E"),interpolate="linear", space ="Lab" )(1000)

anno_col <-list (super_class = c("Organic acids and derivatives" = 'forestgreen',"Nucleosides, nucleotides, and analogues"= 'red2',
                                 "Benzenoids"='orange', 
                                 "Organoheterocyclic compounds"= 'magenta',
                                 "Lipids and lipid-like molecules"='blue', "Organic oxygen compounds"='tan4',
                                 "Organic nitrogen compounds"= 'yellowgreen', "Alkaloids and derivatives"= 'darkgrey',
                                 "Phenylpropanoids and polyketides"= 'mediumvioletred',"Others"='#ECEFF4'))



met_annotation2 <- met_annotation
anno_col2 <- anno_col
colnames(met_annotation2) <- "Metabolite classification                                  Frequency"
names(anno_col2) <- "Metabolite classification                                  Frequency"


frequency_list <- table(met_annotation2[,1])
frequency_list <- frequency_list[rev(order(frequency_list))]
anno_col2$`Metabolite classification                                  Frequency` <- anno_col2$`Metabolite classification                                  Frequency`[names(frequency_list)]

# Indonesia
tbm_IN <- clinmet %>%
  filter (diagnosis_ultimate2 == 'TBM' & cohort == 'Indonesia') %>%
  as.data.frame()

spearman_tbm_IN <- cor(tbm_IN[, met_names], method = 'spearman',
                              use = 'pairwise.complete.obs')


d_mat<-diag(-1,nrow=(length(met_names)+1))
d_mat[(1:length(met_names)),(1:length(met_names))]<-spearman_tbm_IN
colnames(d_mat)<-c(colnames(spearman_tbm_IN),"fake")
rownames(d_mat)<-c(rownames(spearman_tbm_IN),"fake")



p1<- pheatmap(d_mat,  color = myColor, annotation_row = met_annotation2, annotation_colors = anno_col2, main = "Indonesia",show_rownames=F,show_colnames=F,annotation_names_row=F,legend = F,annotation_legend = F)

# Vietnam
tbm_VN <- clinmet %>%
  filter (diagnosis_ultimate2 == 'TBM' & cohort == 'Vietnam') %>%
  as.data.frame()

spearman_tbm_VN <- cor(tbm_VN[, met_names], method = 'spearman',
                              use = 'pairwise.complete.obs')

d_mat<-diag(-1,nrow=(length(met_names)+1))
d_mat[(1:length(met_names)),(1:length(met_names))]<-spearman_tbm_VN
colnames(d_mat)<-c(colnames(spearman_tbm_VN),"fake")
rownames(d_mat)<-c(rownames(spearman_tbm_VN),"fake")

p2<- pheatmap(d_mat,  color = myColor, annotation_row = met_annotation2, annotation_colors = anno_col2, main = "Vietnam",show_rownames=F,show_colnames=F,annotation_names_row=F,legend = T,annotation_legend = T)


tiff(file=paste0(path_output,"/Figure_S1A.tiff"),res=480, units = 'in', width = 15, height = 5,compression = 'lzw')
grid.arrange(p1[[4]],p2[[4]], ncol = 2,widths=c(0.8,1.5))
l = names(anno_col2$`Metabolite classification                                  Frequency`)
grid.text(rev(unname(unlist(frequency_list[l]))),just = "left",
          x = 0.9073,
          y = seq(0, 0.37, (0.37-0)/9)+0.385,gp=gpar(fontsize=10))

grid.lines(x = unit(c(0.706, 0.955), "npc"),
          y = unit(c(0.36, 0.36), "npc"),
          default.units = "npc",
          arrow = NULL, name = NULL,
          gp=gpar(lwd=1.2), draw = TRUE, vp = NULL)
grid.lines(x = unit(c(0.706, 0.955), "npc"),
          y = unit(c(0.82, 0.82), "npc"),
          default.units = "npc",
          arrow = NULL, name = NULL,
          gp=gpar(lwd=1.2), draw = TRUE, vp = NULL)
grid.lines(x = unit(c(0.706, 0.955), "npc"),
          y = unit(c(0.78, 0.78), "npc"),
          default.units = "npc",
          arrow = NULL, name = NULL,
          gp=gpar(lwd=1.2), draw = TRUE, vp = NULL)
dev.off()
```

### Figure S1B:

```{r}
table(clinmet_full$diagnosis_ultimate)
```


```{r}
library(ggrepel)

# Step 1
table(clinmet_full$diagnosis_ultimate, useNA = "always")
clinmet_full_data_S8B<-clinmet_full %>% 
  dplyr::filter(diagnosis_ultimate %in% c('Definite TBM','Probable TBM','Non-infectious control'))%>%
  dplyr::select(c(2:470,478))%>%
  dplyr::mutate(group = ifelse(diagnosis_ultimate %in% c("Definite TBM","Probable TBM") ,"TBM" ,"Control")) #change control to Non-infectious control later !
table(clinmet_full_data_S8B$group)# 54    1067

# step 2: summary
clinmet_full_data_S8B %>%
  tidyr::gather(key = "Metabolites", 
                value = "log2(abundance)", 
                QI09494_C18_neg:TF59_HILIC_neg) %>%
  as.data.frame() -> TBM_mets_long

TBM_mets_long_summary<- TBM_mets_long %>% group_by(Metabolites,group) %>% dplyr::summarise(median_mets=median(`log2(abundance)`,na.rm=T)) %>% ungroup() %>% as.data.frame()
dim(TBM_mets_long_summary)#938   3
table(TBM_mets_long_summary$group)

# step 3: subgroup for each cohort
TBM_mets_long_summary %>%
  dplyr::filter(group == "TBM") %>%
  as.data.frame() -> TBM_mets_long_summary_TBM

TBM_mets_long_summary%>%
  dplyr::filter(group == "Control") %>%
  as.data.frame() -> TBM_mets_long_summary_Control

TBM_mets_wide_summary <- cbind(TBM_mets_long_summary_TBM[,c(1,3)],TBM_mets_long_summary_Control[,3])
names(TBM_mets_wide_summary) <- c("Metabolite", "TBM","Control")

TBM_mets_wide_summary$Differences <- TBM_mets_wide_summary$TBM - TBM_mets_wide_summary$Control


# Step 4: Wilcoxon test to see significantly different metabolites between cohorts (using Wilcoxon because metabolite abundance is not normally distributed)
TBM_mets_pvalue<- lapply(split(TBM_mets_long, TBM_mets_long$Metabolites),function(x) wilcox.test(`log2(abundance)` ~ group, data = x)$p.value)

table((names(TBM_mets_pvalue) == TBM_mets_wide_summary$Metabolite))# 333 TRUE
table((names(TBM_mets_pvalue) %in% TBM_mets_wide_summary$Metabolite))# 469 TRUE

TBM_mets_wide_summary$pvalue<-TBM_mets_pvalue[TBM_mets_wide_summary$Metabolite]

# Step 5: test TF02_C8_pos
TBM_mets_wide_summary %>% dplyr::filter(Metabolite == "TF02_C8_pos")
test<-TBM_mets_long %>% dplyr::filter(Metabolites == "TF02_C8_pos")  

ggplot(test,aes(x= test$group, y = test$"log2(abundance)"))+
  geom_boxplot()+
  stat_compare_means()

# Step 6: FDR
TBM_mets_wide_summary$pvalue <- as.numeric(TBM_mets_wide_summary$pvalue)
TBM_mets_wide_summary$FDR <- p.adjust(TBM_mets_wide_summary$pvalue, method = 'BH')

#Filtering out the most extreme metabolites with a FDR < 1^-10
extreme_pvalues <- TBM_mets_wide_summary %>% dplyr::filter(FDR < 1e-10) %>%as.data.frame()
list_met_extreme_pvalues <- extreme_pvalues$Metabolite

wilcoxon_extreme_met <- TBM_mets_long %>% dplyr::filter(Metabolites %in% list_met_extreme_pvalues) %>%as.data.frame()
dim(wilcoxon_extreme_met)# 344147      4

# step 7: visualization
TBM_mets_wide_summary$Abundance <- 0
TBM_mets_wide_summary$Abundance[TBM_mets_wide_summary$pvalue <0.05]<- TBM_mets_wide_summary$Differences[TBM_mets_wide_summary$pvalue <0.05]
TBM_mets_wide_summary %<>% arrange(desc(pvalue)) 

color_limit <- max(-min(TBM_mets_wide_summary[, "Abundance"], na.rm = TRUE), max(TBM_mets_wide_summary[, "Abundance"], na.rm = TRUE), na.rm=TRUE)*1.01

colour_scheme_outcome_L <- c("blue", "cornflowerblue")
colour_scheme_outcome_R <- c("tomato", "red")

Metabolites <- c(
  "hydroxyisocaproate",
  "3-hydroxyoctanoate",
  "hydroxyisobutyrate",
  "C4-OH carnitine",
  "tryptophan",
  "p-hydroxyphenylacetate",
  "phenyllactate",
  "N-carbamoyl-beta-alanine",
  "N6,N6,N6-trimethyllysine",
  "sebacate")

rename_metabolites <- list("hydroxyisocaproate" = "FA 6:0;OH",
                           "3-hydroxyoctanoate"= "FA 8:0;3OH",
                           "hydroxyisobutyrate" = "FA 4:0;OH",
                           "C4-OH carnitine" = "CAR 4:0;OH",
                           "tryptophan" = "Tryptophan",
                           "p-hydroxyphenylacetate"="p-Hydroxyphenylacetate",
                           "phenyllactate" = "Phenyllactate",
                           "N-carbamoyl-beta-alanine" = "N-carbamoyl-beta-alanine",
                           "N6,N6,N6-trimethyllysine" = "N6,N6,N6-Trimethyllysine",
                           "sebacate" = "Sebacate")

color_top_hit <- TBM_mets_wide_summary$Metabolite %in% filter(names_ID,Metabolite %in% Metabolites)$Metabol_names

options(ggrepel.max.overlaps = Inf)

tmp <- TBM_mets_wide_summary
tmp$Metabolite <- names_ID$Metabolite[match(TBM_mets_wide_summary$Metabolite, names_ID$Metabol_names)]
rename_metabolites[tmp$Metabolite]

x_logticks <- annotation_logticks(sides = "b", short = unit(0.05, "cm"), mid = unit(0.1, "cm"), long = unit(0.15, "cm"))

p<-ggplot(data=tmp)+
  geom_point(aes(x=2**Differences, 
                 y=pvalue,
                 fill=Abundance),
             shape=21, stroke = 0,size=2.5,
             alpha=1,
             )+
  
  geom_point(data = tmp %>% filter(Metabolite %in% filter(names_ID,Metabolite %in% Metabolites)$Metabolite), 
             aes(x = 2**Differences, y = pvalue), color = "black",shape=21, fill="#A2F001",size=2.5)+
  #scale_color_manual(values = c("black", "green")) +
  scale_fill_gradientn(guide = "colourbar", 
                       colours = c(colour_scheme_outcome_L, "lightgrey", colour_scheme_outcome_R),
                       breaks=c(-2,2),
                       labels=c("Higher in NIC","Higher in TBM"), 
                       limits = c(-color_limit, color_limit)) +
  
  labs(fill="Metabolite \nabundance    ",space="Lab")+#scale_x_log2(limits = 2**c(-5,5), breaks = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8, 16,32), labels = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8,16,32), name = "Ratio Vietnam vs. Indonesia")+
  scale_y_continuous(trans=reverselog_trans(10), 
                     breaks = trans_breaks("log10", function(x) 10^x), 
                     labels = trans_format("log10", math_format(10^.x)), 
                     name = "Unadjusted p-value")+
  
  scale_x_continuous(trans=log2_trans(), 
                     breaks = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8, 16,32), 
                     labels = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8,16,32),
                     name = "Ratio TBM vs. Non-infectious control")+
  x_logticks +
  
  geom_vline(xintercept=1, color='black',linewidth=0.2)+
  geom_text_repel(max.overlaps = Inf,
                  vjust = 0.52,
                  hjust = 0.52,
                  nudge_x = -0.01,
                  box.padding = 0.5,
                  direction = 'both',
                  xlim = c(-Inf, Inf),
                  ylim = c(-2, Inf),
                  lineheight = 1.1,
                  min.segment.length = 0,
                  size=3,
                  aes(x=2**Differences,
                      y= pvalue,
                      label=ifelse(Metabolite %in% filter(names_ID,Metabolite %in% Metabolites)$Metabolite,
                                   rename_metabolites[as.character(Metabolite)],''))) +
  theme_classic()+
  theme(text = element_text(size = 9))+
  theme(legend.position="bottom")+
  theme(legend.title = element_text(face="bold"))+coord_cartesian(xlim=c(1/4,4))+
  theme(axis.title.x = element_text(face="bold", colour="black", size=15),
            axis.title.y = element_text(face="bold", colour="black", size=15),
            axis.text.y = element_text( colour="black", size=15),
            axis.text.x = element_text( colour="black", size=13),
            axis.ticks.x=element_blank(),
            legend.text = element_text(size = 14),
            legend.title = element_text( colour="black", size=15),
            legend.key = element_rect(colour = NA),
            strip.text=element_text(face = "bold",size=15))

panel_height <- unit(0.5,"npc") - sum(ggplotGrob(p)[["widths"]][-3]) - unit(1,"line")
p <- p + guides(fill= guide_colorbar(barwidth=panel_height))

tiff(file=paste0(path_output,"/Figure_S1B.tiff"),res=600, units = 'in', width =8, height =8,compression = 'lzw')
p
dev.off()



#export for Edwin to perform pathway analysis
TBM_mets_wide_summary$Metabolite_name <- names_ID$Metabolite[match(TBM_mets_wide_summary$Metabolite, names_ID$Metabol_names)]

TBM_mets_wide_summary <- TBM_mets_wide_summary[,c('Metabolite','Metabolite_name','TBM','Control','Differences','pvalue','FDR','Abundance')]

save(TBM_mets_wide_summary,file="metabolites_TBMvsNIC_stats.Rdata")


```

check data: volcano plot TBM vs Other CNS

```{r}
library(ggrepel)

# Step 1
table(clinmet_full$diagnosis_ultimate, useNA = "always")
clinmet_full_data_S8B<-clinmet_full %>% 
  dplyr::filter(diagnosis_ultimate %in% c('Definite TBM','Probable TBM','Bacterial meningitis','Cryptococcal meningitis'))%>%
  dplyr::select(c(2:470,478))%>%
  dplyr::mutate(group = ifelse(diagnosis_ultimate %in% c("Definite TBM","Probable TBM") ,"TBM" ,"Control")) #change control to Non-infectious control later !
table(clinmet_full_data_S8B$group)# 112    1067

# step 2: summary
clinmet_full_data_S8B %>%
  tidyr::gather(key = "Metabolites", 
                value = "log2(abundance)", 
                QI09494_C18_neg:TF59_HILIC_neg) %>%
  as.data.frame() -> TBM_mets_long

TBM_mets_long_summary<- TBM_mets_long %>% group_by(Metabolites,group) %>% dplyr::summarise(median_mets=median(`log2(abundance)`,na.rm=T)) %>% ungroup() %>% as.data.frame()

dim(TBM_mets_long_summary)#938   3
table(TBM_mets_long_summary$group)

# step 3: subgroup for each cohort
TBM_mets_long_summary %>%
  dplyr::filter(group == "TBM") %>%
  as.data.frame() -> TBM_mets_long_summary_TBM

TBM_mets_long_summary%>%
  dplyr::filter(group == "Control") %>%
  as.data.frame() -> TBM_mets_long_summary_Control

TBM_mets_wide_summary <- cbind(TBM_mets_long_summary_TBM[,c(1,3)],TBM_mets_long_summary_Control[,3])
names(TBM_mets_wide_summary) <- c("Metabolite", "TBM","Control")

TBM_mets_wide_summary$Differences <- TBM_mets_wide_summary$TBM - TBM_mets_wide_summary$Control


# Step 4: Wilcoxon test to see significantly different metabolites between cohorts (using Wilcoxon because metabolite abundance is not normally distributed)
TBM_mets_pvalue<- lapply(split(TBM_mets_long, TBM_mets_long$Metabolites),function(x) wilcox.test(`log2(abundance)` ~ group, data = x)$p.value)

table((names(TBM_mets_pvalue) == TBM_mets_wide_summary$Metabolite))# 333 TRUE
table((names(TBM_mets_pvalue) %in% TBM_mets_wide_summary$Metabolite))# 469 TRUE

TBM_mets_wide_summary$pvalue<-TBM_mets_pvalue[TBM_mets_wide_summary$Metabolite]

# Step 5: test TF02_C8_pos
TBM_mets_wide_summary %>% dplyr::filter(Metabolite == "TF02_C8_pos")
test<-TBM_mets_long %>% dplyr::filter(Metabolites == "TF02_C8_pos")  

ggplot(test,aes(x= test$group, y = test$"log2(abundance)"))+
  geom_boxplot()+
  stat_compare_means()

# Step 6: FDR
TBM_mets_wide_summary$pvalue <- as.numeric(TBM_mets_wide_summary$pvalue)
TBM_mets_wide_summary$FDR <- p.adjust(TBM_mets_wide_summary$pvalue, method = 'BH')

#Filtering out the most extreme metabolites with a FDR < 1^-10
extreme_pvalues <- TBM_mets_wide_summary %>% dplyr::filter(FDR < 1e-10) %>%as.data.frame()
list_met_extreme_pvalues <- extreme_pvalues$Metabolite

wilcoxon_extreme_met <- TBM_mets_long %>% dplyr::filter(Metabolites %in% list_met_extreme_pvalues) %>%as.data.frame()
dim(wilcoxon_extreme_met)# 344147      4

# step 7: visualization
TBM_mets_wide_summary$Abundance <- 0
TBM_mets_wide_summary$Abundance[TBM_mets_wide_summary$pvalue <0.05]<- TBM_mets_wide_summary$Differences[TBM_mets_wide_summary$pvalue <0.05]
TBM_mets_wide_summary %<>% arrange(desc(pvalue))

color_limit <- max(-min(TBM_mets_wide_summary[,"Abundance"], na.rm = TRUE), max(TBM_mets_wide_summary[, "Abundance"], na.rm = TRUE), na.rm=TRUE)*1.01

colour_scheme_outcome_L <- c("blue", "cornflowerblue")
colour_scheme_outcome_R <- c("tomato", "red")

Metabolites <- c(
  "hydroxyisocaproate",
  "3-hydroxyoctanoate",
  "hydroxyisobutyrate",
  "C4-OH carnitine",
  "tryptophan",
  "p-hydroxyphenylacetate",
  "phenyllactate",
  "N-carbamoyl-beta-alanine",
  "N6,N6,N6-trimethyllysine",
  "sebacate")

rename_metabolites <- list("hydroxyisocaproate" = "FA 6:0;OH",
                           "3-hydroxyoctanoate"= "FA 8:0;3OH",
                           "hydroxyisobutyrate" = "FA 4:0;OH",
                           "C4-OH carnitine" = "CAR 4:0;OH",
                           "tryptophan" = "Tryptophan",
                           "p-hydroxyphenylacetate"="p-Hydroxyphenylacetate",
                           "phenyllactate" = "Phenyllactate",
                           "N-carbamoyl-beta-alanine" = "N-carbamoyl-beta-alanine",
                           "N6,N6,N6-trimethyllysine" = "N6,N6,N6-Trimethyllysine",
                           "sebacate" = "Sebacate")

color_top_hit <- TBM_mets_wide_summary$Metabolite %in% filter(names_ID,Metabolite %in% Metabolites)$Metabol_names

options(ggrepel.max.overlaps = Inf)

tmp <- TBM_mets_wide_summary
tmp$Metabolite <- names_ID$Metabolite[match(TBM_mets_wide_summary$Metabolite, names_ID$Metabol_names)]
rename_metabolites[tmp$Metabolite]

x_logticks <- annotation_logticks(sides = "b", short = unit(0.05, "cm"), mid = unit(0.1, "cm"), long = unit(0.15, "cm"))

p<-ggplot(data=tmp)+
  geom_point(aes(x=2**Differences, 
                 y=pvalue,
                 fill=Abundance),
             shape=21, stroke = 0,size=2.5,
             alpha=1,
             )+
  
  # geom_point(data = tmp %>% filter(Metabolite %in% filter(names_ID,Metabolite %in% Metabolites)$Metabolite), 
  #            aes(x = 2**Differences, y = pvalue), color = "black",shape=21, fill="#A2F001",size=2.5)+
  #scale_color_manual(values = c("black", "green")) +
  scale_fill_gradientn(guide = "colourbar", 
                       colours = c(colour_scheme_outcome_L, "lightgrey", colour_scheme_outcome_R),
                       breaks=c(-1,1),
                       labels=c("Higher in Other CNS","Higher in TBM"), 
                       limits = c(-color_limit, color_limit)) +
  
  labs(fill="Metabolite \nabundance    ",space="Lab")+#scale_x_log2(limits = 2**c(-5,5), breaks = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8, 16,32), labels = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8,16,32), name = "Ratio Vietnam vs. Indonesia")+
  scale_y_continuous(trans=reverselog_trans(10), 
                     breaks = trans_breaks("log10", function(x) 10^x), 
                     labels = trans_format("log10", math_format(10^.x)), 
                     name = "Unadjusted p-value")+
  
  scale_x_continuous(trans=log2_trans(), 
                     breaks = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8, 16,32), 
                     labels = c(1/32, 1/16, 1/8,1/4,1/2, 1,2,4,8,16,32),
                     name = "Ratio TBM vs. Other CNS Infection")+
  x_logticks +
  
  geom_vline(xintercept=1, color='black',linewidth=0.2)+
  geom_hline(yintercept=0.05, color='black',linewidth=0.2,linetype='dashed')+
  geom_vline(xintercept=2, color='black',linewidth=0.2,linetype='dashed')+
  geom_vline(xintercept=0.5, color='black',linewidth=0.2,linetype='dashed')+
  theme_classic()+
  theme(text = element_text(size = 9))+
  theme(legend.position="bottom")+
  theme(legend.title = element_text(face="bold"))+coord_cartesian(xlim=c(1/4,4))+
  theme(axis.title.x = element_text(face="bold", colour="black", size=15),
            axis.title.y = element_text(face="bold", colour="black", size=15),
            axis.text.y = element_text( colour="black", size=15),
            axis.text.x = element_text( colour="black", size=13),
            axis.ticks.x=element_blank(),
            legend.text = element_text(size = 14),
            legend.title = element_text( colour="black", size=15),
            legend.key = element_rect(colour = NA),
            strip.text=element_text(face = "bold",size=15)) + guides(fill= guide_colorbar(barwidth=5))

panel_height <- unit(0.5,"npc") - sum(ggplotGrob(p)[["widths"]][-3]) - unit(0.5,"line")
p <- p + guides(fill= guide_colorbar(barwidth=panel_height))
p


tiff(file=paste0(path_output,"/Figure_Volcano_TBMvsOtherCNS.tiff"),res=600, units = 'in', width =12, height =12,compression = 'lzw')
p
dev.off()


met_log2fc_TBMvsOtherCNS <- tmp
save(met_log2fc_TBMvsOtherCNS,file='met_log2fc_TBMvsOtherCNS.Rdata')


clinmet_full_TBM_OCNS<-clinmet_full %>% 
  dplyr::filter(diagnosis_ultimate %in% c('Definite TBM','Probable TBM','Bacterial meningitis','Cryptococcal meningitis'))%>%
  dplyr::select(c(1:470,478))%>%
  dplyr::mutate(group = ifelse(diagnosis_ultimate %in% c("Definite TBM","Probable TBM") ,"TBM" ,"OCNS")) #change control to Non-infectious control later !
table(clinmet_full_TBM_OCNS$group) #112 1067
save(clinmet_full_TBM_OCNS,file='clinmet_full_TBM_OCNS.Rdata')

```

### Figure S1C:

```{r}
library("factoextra")
clinmet_full_data <- clinmet_full 
clinmet_full_data <- clinmet_full_data[complete.cases(clinmet_full_data[met_names]),]
pca_indo <- prcomp(clinmet_full_data[, met_names], center = T, scale. = T)

table(clinmet_full_data$diag_combine)# 54                    110                   1062
table(clinmet_full_data$diagnosis_ultimate)# 49  , 61,  682, 54  ,  380 

clinmet_full_data_S8<-clinmet_full_data %>% mutate(
  diag_combine_2 = case_when( diag_combine == "Other CNS infection" & 
                                diagnosis_ultimate == "Bacterial meningitis" ~ "Bacterial meningitis",
                              
                               diag_combine == "Other CNS infection" & 
                                diagnosis_ultimate == "Cryptococcal meningitis" ~ "Cryptococcal meningitis",
                              
                              diag_combine == "Non-infectious control" ~ "Non-infectious control",
                               diag_combine == "TBM" ~ "TBM"
                                ))

clinmet_full_data_S8 %<>% mutate(diag_combine_2 = factor(diag_combine_2,levels = c('Non-infectious control',
                                                                                   'Bacterial meningitis',
                                                                                   'Cryptococcal meningitis',
                                                                                   'TBM'
                                                                                   
                                                                                   
                                                                                   )))

table(clinmet_full_data_S8$diag_combine_2, useNA = "always")

p1 <- fviz_pca_ind(pca_indo, geom.ind = "point", pointshape = 21, 
                   pointsize = 5, alpha=0.7,
                   fill.ind = clinmet_full_data_S8$diag_combine_2, 
                   col.ind = "grey50", 
                   palette = c("#bebada", "#ebbb8b", "#A3BE8C", "#81A1C1"),
                   addEllipses = TRUE,
                   label = "var",
                   col.var = "black",
                   repel = TRUE,
                   legend.title = "Etiology",
                   title="")+ theme_classic()+
  theme(text = element_text(size = 9))+
  theme(legend.position="bottom")+
  labs(x = "PC1 (41%)", y = "PC2 (7.2%)")+
  theme(plot.title = element_text(face="bold",size=9))+
  theme(legend.title = element_text(face="bold")) +
  theme(axis.title.x = element_text(face="bold", colour="black", size=15),
        axis.title.y = element_text(face="bold", colour="black", size=15),
        axis.text.y = element_text( colour="black", size=15),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 14),
        legend.title = element_text( colour="black", size=15),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=15))+
  scale_color_manual(values=c("#bebada", "#ebbb8b", "#A3BE8C", "#81A1C1"))+
  scale_fill_manual(values=c("#bebada", "#ebbb8b", "#A3BE8C", "#81A1C1"))

#Export
tiff(file=paste0(path_output,"/Figure_S1C.tiff"),res=450, units = 'in', width = 10, height =10,compression = 'lzw')
p1
dev.off()

```

### Figure S1D:

```{r}

print(load(file="Derive_Results/IN_correlation_learning_set_2022_10_12.Rdata"))
Metabol_names<-colnames(branks)
branks<-as.matrix(branks)
branks_sum<-matrixStats::colQuantiles(branks,probs = c(0.025,0.975,0.5)) %>% as.data.frame()
branks_sum$means_ranks<-colMeans(branks)

#branks_sum<- as.data.frame(branks_sum) %>% arrange(means_ranks)
branks_sum<- as.data.frame(branks_sum)
branks_sum<- branks_sum[order(branks_sum$means_ranks), , drop = FALSE]


names(branks_sum)[1:3]<-paste0("rank-",colnames(branks_sum[,1:3]))
branks_sum$Metabol_names<-rownames(branks_sum)

dat<-branks_sum %>% mutate(name = reorder(Metabol_names, desc(`rank-50%`)))
branks_sum_IN<-branks_sum

print(load(file="Derive_Results/HRs_IN.Rdata"))


dat_HR_IN<-merge(branks_sum_IN,HRs,by="Metabol_names") %>% arrange((`rank-50%`))
dat_HR_IN$select<-ifelse(((dat_HR_IN[,"rank-2.5%"]<=75 & dat_HR_IN[,"rank-97.5%"]<=225)|
                           (dat_HR_IN[,"rank-2.5%"]>=244 & dat_HR_IN[,"rank-97.5%"]>=394))& dat_HR_IN$significant_screening=="yes","yes","no")

table(dat_HR_IN$select) #361 108: 107 + 1 unknown compound

dat.plot_IN<-dat_HR_IN %>% filter(select=="yes") %>% arrange(desc(`rank-50%`))
Metabol_tophit_names_IN<-dat.plot_IN$Metabol_names

dim(branks_sum)# 469   5
dat<-branks_sum %>%
  mutate(name = reorder(Metabol_names, desc(`rank-50%`)),
         Selection=ifelse(Metabol_names%in%Metabol_tophit_names_IN,
                       ifelse(means_ranks>225,"Protective","Harmful"),"un-selected"))

summary(branks_sum$means_ranks)

dat<-branks_sum %>%
  mutate(name = reorder(Metabol_names, desc(`rank-50%`)),
         Selection=case_when(
         Metabol_names %in% dat.plot_IN$Metabol_names & Metabol_names != "TF21_HILIC_pos" & branks_sum$"rank-2.5%"  >=244 &  branks_sum$"rank-97.5%">=394~ "Protective",
         Metabol_names %in% dat.plot_IN$Metabol_names & Metabol_names != "TF21_HILIC_pos" & branks_sum$"rank-2.5%"  <=75 &  branks_sum$"rank-97.5%"<=225  ~ "Harmful",
         T ~ "un-selected"
         ))

table(dat$Selection)#         101           6         362 tophit

dat%>% 
  ggplot( aes(x=name, y=`rank-50%`,group=Selection,color=Selection)) +
    geom_point() +
  geom_errorbar(aes(ymin = `rank-2.5%`, ymax = `rank-97.5%`),alpha=0.5)+
  #scale_color_manual(values=c("#de2d26","#2ca25f","#bcbddc"))+
 scale_color_manual(values=c("#A50026","#313695","#bdbdbd"))+
  scale_fill_manual(values=c("#A50026","#313695","#bdbdbd"))+
  theme(axis.text.y = element_blank())+
  geom_hline(yintercept =c(75,225,244,394), linetype = 'dashed', color= 'red',size=1)+
    coord_flip() +
    xlab("") +
    theme_bw()+ theme(legend.position="bottom")+ylab("Distribution ranking of  metabolite-mortality associations")+
  
  theme(panel.grid.minor = element_blank(),
        axis.title.x = element_text(face="bold", colour="black", size=15),
        axis.title.y = element_text(face="bold", colour="black", size=15),
        axis.text.y = element_blank(),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=12),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=12))+ggtitle("Bi-directional ranking
") +  theme(plot.title = element_text(hjust = 0.5,face="bold", colour="black", size=20))->p


#Export
tiff(file=paste0(path_output,"/Figure_S1D.tiff"),res=480, units = 'in', width = 10, height =10,compression = 'lzw')
p
dev.off()

```

### Figure S1E:

```{r}

dat.plot_IN<-dat_HR_IN %>% filter(select=="yes") %>% arrange(desc(`rank-50%`))
Metabol_tophit_names_IN<-dat.plot_IN$Metabol_names

names_ID<-subset(untargeted_annotated,select=c("CompoundID","Metabolite"))
names_ID$Metabol_names<-gsub("-", "_", names_ID$CompoundID, fixed = TRUE)

dat.plot_IN<-merge(dat.plot_IN,names_ID,by="Metabol_names") 
dat.plot_IN%<>% filter(Metabolite!='Unknown_of_interest')

dat.plot_IN %<>% mutate(Metabolite = plyr::mapvalues(dat.plot_IN$Metabolite,
                                                     from = names(rename_metabolites),
                                                     to = unname(unlist(rename_metabolites))))

p1<-dat.plot_IN %>% mutate(name = reorder(Metabolite, desc(`rank-50%`))) %>%
  ggplot( aes(x=name, y=`rank-50%`)) +
    geom_point(aes(col="red"),size=2) +
  geom_errorbar(aes(ymin = `rank-2.5%`, ymax = `rank-97.5%`,alpha=0.01),size=1)+
  theme_classic()+
  theme(axis.text.y = element_blank())+
  geom_hline(yintercept =c(75,225,244,394), linetype = 'dashed', color= 'red',size=1)+
    coord_flip() +
    xlab("") +ylab("Rank on HR")+
  ggtitle("Top-hit ranking \n (Indonesia)")+
    theme_bw()+ theme(legend.position="none")+theme(panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=13),
        axis.title.y = element_text(face="bold", colour="black", size=13),
        axis.text.y = element_text(face="bold", colour="black", size=13),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=13),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=18),
        plot.title = element_text(face = "bold",size=18,hjust = 0.5))

p2<-dat.plot_IN%>% mutate(name = reorder(Metabolite, desc(`rank-50%`))) %>%
  ggplot( aes(x=name, y=log.HR.est_screening)) +
    geom_point(aes(col=significant_screening),size=2)+
  scale_color_manual(values=c("#116400","#001164"))+
   scale_color_manual(values=c("grey40","#001164"))+
  scale_y_continuous(breaks=log(c(1/256,1/32,1/8,1/2,1,2,4,8)),labels=c(expression(1/256),expression(1/32),expression(1/8),expression(1/2),1,2,4,8))+
  geom_errorbar(aes(ymin = log.HR.low_screening, ymax =log.HR.hi_screening,alpha=0.01,col=significant_screening),size=1)+
  theme_classic()+
  theme(axis.text.y = element_blank())+
  geom_hline(yintercept =c(0), linetype = 'dashed', color= '#001164',size=1) +
      coord_flip() +
    xlab("") +ylab("Hazard ratio")+
  ggtitle("Screening cohort \n (Indonesia)")+
    theme_bw()+ theme(legend.position="none")+theme(panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=13),
        axis.title.y = element_text(face="bold", colour="black", size=13),
        axis.text.y = element_text(face="bold", colour="black", size=13),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=13),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=13),
        plot.title = element_text(face = "bold",size=18,hjust = 0.5))

p3<-dat.plot_IN%>% mutate(name = reorder(Metabolite, desc(`rank-50%`))) %>%
  ggplot( aes(x=name, y=log.HR.est_wc_validation)) +
    geom_point(aes(col=significant_wc_validation),size=2)+scale_y_continuous(breaks=log(c(1/32,1/8,1/2,1,2,4,8)),labels=c(expression(1/32),expression(1/8),expression(1/2),1,2,4,8))+
  geom_errorbar(aes(ymin = log.HR.low_wc_validation, ymax =log.HR.hi_wc_validation,alpha=0.01,col=significant_wc_validation),size=1)+
  scale_color_manual(values=c("#116400","#001164"))+
  theme_classic()+
  theme(axis.text.y = element_blank())+
  geom_hline(yintercept =c(0), linetype = 'dashed', color= "#001164",size=1)+
      coord_flip() +
    xlab("") +ylab("Hazard ratio")+
  ggtitle("Within-cohort validation \n (Indonesia)")+
    theme_bw()+ theme(legend.position="none")+theme(panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=13),
        axis.title.y = element_text(face="bold", colour="black", size=13),
        axis.text.y = element_blank(),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=13),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=13),
        plot.title = element_text(face = "bold",size=18,hjust = 0.5))


p4<-dat.plot_IN%>% mutate(name = reorder(Metabolite, desc(`rank-50%`))) %>%
  ggplot( aes(x=name, y=log.HR.est_VN_validation)) +
    geom_point(aes(col=significant_external_validation),size=2)+scale_y_continuous(breaks=log(c(1/32,1/8,1/2,1,2,4,8)),labels=c(expression(1/32),expression(1/8),expression(1/2),1,2,4,8))+
  geom_errorbar(aes(ymin = log.HR.low_VN_validation, ymax =log.HR.hi_VN_validation,alpha=0.01,col=significant_external_validation),size=1)+
  scale_color_manual(values=c("#116400","#001164"))+
  theme_classic()+
  theme(axis.text.y = element_blank())+
  geom_hline(yintercept =c(0), linetype = 'dashed', color= "#001164",size=1)+
      coord_flip() +
    xlab("") +ylab("Hazard ratio")+
  ggtitle("External validation \n (Vietnam)")+
    theme_bw()+ theme(legend.position="none")+theme(panel.grid.minor = element_blank(),axis.title.x = element_text(face="bold", colour="black", size=13),
        axis.title.y = element_text(face="bold", colour="black", size=13),
        axis.text.y = element_blank(),
        axis.text.x = element_text( colour="black", size=13),
        axis.ticks.x=element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text( colour="black", size=13),
        legend.key = element_rect(colour = NA),
        strip.text=element_text(face = "bold",size=13),
        plot.title = element_text(face = "bold",size=18,hjust = 0.5))

tiff(file=paste0(path_output,"/Figure_S1E_long.tiff"),res=480, units = 'in', width = 17, height =30,compression = 'lzw')
cowplot::plot_grid(p2,p3,p4, nrow = 1,rel_widths = c(0.4, 0.18,0.18))
dev.off()

```
